### 2018-07-10 submission/manuscript.Rmd
fix up discussion first paragraph

Previous studies using regular DNA polymerase have reported GC dependent amplification differences [@polz_bias_1998] and we confirm that these differences still persist in high fidelity DNA polymerase when analyzing mock community samples [Figure 1]. Additionally, these specific polymerases also have different sequence error rate and chimera prevalence in these samples [Figure 2]. Chimera prevalence in particular was strongly associated with the total number of OTUs detected in a sample [Figure 2E]. The total number of OTUs also varied by polymerase used for mock community samples [Fiugre 3A]. The range in the number of OTUs observed increased with the number of cycles and the change in range was dependent on the polymerase for mock community and fecal samples [Figure 3]. However, these differences did not translate to large bacterial community composition changes based on polymerase [Figure 4]. Even though there were bacterial community composition changes based on the number of cycles they occured independently of the polymerase used [Figure 4]. Additionally, there were no differences observed within the number of cycles between any of the polymerases for fecal samples. When Random Forest models were created to classify number of cycles, polymerase, and individual based on the 16S rRNA gene sequencing the model performance for individuals was significantly greater than all other models tested [Figure 5A]. In addition to the better model performance of individuals the most important OTUs to this model were not the most important OTUs in the other models [Figure 5B]. Our observations show that differences between high fidelity DNA polymerases can be found. However, when using 16S rRNA gene sequencing these technical differences are dwarfed by biological differences between individuals.

>>>>

Previous studies have reported GC dependent amplification differences [@polz_bias_1998] and we confirm that these differences still persist in high fidelity DNA polymerase when analyzing mock community samples [Figure 1]. Additionally, these specific polymerases have different sequence error rates and chimera prevalence [Figure 2]. Chimera prevalence in particular was strongly associated with the total number of OTUs detected in a mock community [Figure 2E]. This total number of OTUs also varied by polymerase used for these mock community samples [Fiugre 3A]. The range in the number of OTUs observed increased with the number of cycles and the change in range was dependent on the polymerase for both mock community and fecal samples [Figure 3]. However, these differences did not translate to large bacterial community composition changes based on polymerase [Figure 4]. Even though there were bacterial community composition changes based on the number of cycles they occured independently of the polymerase used [Figure 4]. Additionally, there were no differences observed within the number of cycles between any of the polymerases for fecal samples. When Random Forest models were created to classify number of cycles, polymerase, and individual based on the 16S rRNA gene sequencing, the model performance for individuals was significantly greater than all other models tested [Figure 5A]. In addition to the better model performance of individuals, the most important OTUs to this model were not the most important OTUs for the other models tested [Figure 5B]. Our observations show that differences between high fidelity DNA polymerases can be found. However, when using 16S rRNA gene sequencing these technical differences are dwarfed by biological differences between individuals.

### 2018-07-10 submission/manuscript.tex
fix up discussion first paragraph

Previous studies using regular DNA polymerase have reported GC dependent
amplification differences (Polz \& Cavanaugh, 1998) and we confirm that
these differences still persist in high fidelity DNA polymerase when
analyzing mock community samples {[}Figure 1{]}. Additionally, these
specific polymerases also have different sequence error rate and chimera
prevalence in these samples {[}Figure 2{]}. Chimera prevalence in
particular was strongly associated with the total number of OTUs
detected in a sample {[}Figure 2E{]}. The total number of OTUs also
varied by polymerase used for mock community samples {[}Fiugre 3A{]}.
The range in the number of OTUs observed increased with the number of
cycles and the change in range was dependent on the polymerase for mock
community and fecal samples {[}Figure 3{]}. However, these differences
did not translate to large bacterial community composition changes based
on polymerase {[}Figure 4{]}. Even though there were bacterial community
composition changes based on the number of cycles they occured
independently of the polymerase used {[}Figure 4{]}. Additionally, there
were no differences observed within the number of cycles between any of
the polymerases for fecal samples. When Random Forest models were
created to classify number of cycles, polymerase, and individual based
on the 16S rRNA gene sequencing the model performance for individuals
was significantly greater than all other models tested {[}Figure 5A{]}.
In addition to the better model performance of individuals the most
important OTUs to this model were not the most important OTUs in the
other models {[}Figure 5B{]}. Our observations show that differences
between high fidelity DNA polymerases can be found. However, when using
16S rRNA gene sequencing these technical differences are dwarfed by
biological differences between individuals.

>>>>

Previous studies have reported GC dependent amplification differences
(Polz \& Cavanaugh, 1998) and we confirm that these differences still
persist in high fidelity DNA polymerase when analyzing mock community
samples {[}Figure 1{]}. Additionally, these specific polymerases have
different sequence error rates and chimera prevalence {[}Figure 2{]}.
Chimera prevalence in particular was strongly associated with the total
number of OTUs detected in a mock community {[}Figure 2E{]}. This total
number of OTUs also varied by polymerase used for these mock community
samples {[}Fiugre 3A{]}. The range in the number of OTUs observed
increased with the number of cycles and the change in range was
dependent on the polymerase for both mock community and fecal samples
{[}Figure 3{]}. However, these differences did not translate to large
bacterial community composition changes based on polymerase {[}Figure
4{]}. Even though there were bacterial community composition changes
based on the number of cycles they occured independently of the
polymerase used {[}Figure 4{]}. Additionally, there were no differences
observed within the number of cycles between any of the polymerases for
fecal samples. When Random Forest models were created to classify number
of cycles, polymerase, and individual based on the 16S rRNA gene
sequencing, the model performance for individuals was significantly
greater than all other models tested {[}Figure 5A{]}. In addition to the
better model performance of individuals, the most important OTUs to this
model were not the most important OTUs for the other models tested
{[}Figure 5B{]}. Our observations show that differences between high
fidelity DNA polymerases can be found. However, when using 16S rRNA gene
sequencing these technical differences are dwarfed by biological
differences between individuals.
